---
title: "Lab 3 Homework"
author: "Ritwik Varma"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: spacelab
    toc: no
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Mammals Sleep
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R.
```{r}
str(mammals_sleep_allison_cicchetti_1976)
```
This data is taken from allison cicchetti 1976

2. Store these data into a new data frame `sleep`.
```{r}
sleep <- mammals_sleep_allison_cicchetti_1976
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  
```{r}
dim(sleep)
```
62 observations and 11 variables 

4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
table(sleep==-999)
```
There are 38 numbers that aren't real data, but none of them are represented as NAs as shown by the above function which returned TRUE for 38 values. 



5. Show a list of the column names is this data frame.
```{r}
colnames(sleep)
```

6. How many herbivores are represented in the data?  
```{r}
sum(sleep[9]==5)
```
There are 14 herbivores 

7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 1kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
large <- subset(sleep, sleep[2]>=200)  
small <- subset(sleep, sleep[2]<=1)
```

8. What is the mean weight for both the small and large mammals?
```{r}
large_weight <- large$`body weight in kg`
small_weight <- small$`body weight in kg`
mean(large_weight)
mean(small_weight)
```
Mean weight for small mammals is 0.33kg 
Mean weight for large animals is 1596 kg 


9. Using a similar approach as above, do large or small animals sleep longer on average?  
```{r}
large_sleep <- large$`total sleep (hrs/day)  (sum of slow wave and paradoxical sleep)`
y <- large_sleep[-c(4,7)]
mean(y)
```

```{r}
small_sleep <- small$`total sleep (hrs/day)  (sum of slow wave and paradoxical sleep)`
mean(small_sleep)

```
On average small animals sleep longer. 


10. Which animal is the sleepiest among the entire dataframe?
```{r}
sleepiest <- max(sleep$`total sleep (hrs/day)  (sum of slow wave and paradoxical sleep)`)
print(sleep[33,])
```
The Brown Bat is the sleepiest animal in the data frame
```{r}

```

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   